---
title: "Plots for manuscript"
author: "Tory Madden"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output: 
  html_document:
    theme: yeti
    hightlight: pygments
    toc: true
    toc_depth: 5
    toc_float: true
    code_folding: show
---

```{r setup, include = FALSE}
# Load packages
library(readr)
library(tidyverse)
library(dplyr)
library(tidyr)
library(magrittr)
library(RColorBrewer)
library(ggplot2)

# knitr setup
knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE,
                      fig.height = 8,
                      fig.width = 8,
                      fig.path = './figures/manuscript_plots/')
```

```{r data, include = TRUE}

# Import data

df_exp <- read_csv('/Users/tory/Google Drive/Git/CC_threshold/Juliane_all_data/data/R_data_complete_expectancies')
```

# Plot expectancy ratings

## Figure 3a
Produce single plot showing expectancy ratings by CS across the 3 phases of the experiment.  Use different colours and shapes for each CS and jitter the position of the error bars and their corresponding points and lines, so as to make them easier to distinguish.

```{r plot, echo = TRUE}

df_exp %<>% mutate(phase = case_when(
    block == "BL_acq1" ~ "baseline",
    block == "acq1_2" | block == "acq2_3" | block == "acq_test" ~ "acquisition",
    block == "test1_2" | block == "test2_3" | block == "after_test3"  ~ "test"))

df_exp_plot <- df_exp %>% 
    group_by(phase, CS) %>%
    summarise(mean = mean(rating, na.rm = TRUE),
              se = sd(rating, na.rm = TRUE)/sqrt(n()))
    
df_exp_plot$phase <- factor(df_exp_plot$phase, levels = c("baseline", "acquisition", "test"))

Figure_3a <- ggplot(data = df_exp_plot) +
    aes(x = phase,
        y = mean,
        group = CS,
        colour = CS) +
    geom_point(position = position_dodge(0.1),
               aes(shape = CS),
               size = 3) +
    scale_color_manual(name = "CS", values=c("blue", "red")) +
    geom_errorbar(aes(ymin = mean-se,
                      ymax = mean+se),
                  width = 0.2,
                  position = position_dodge(0.1)) +
    geom_line(position = position_dodge(0.1)) +
    scale_shape_discrete(solid = FALSE) +
    scale_y_continuous(limits = c(0,10),
                       breaks = c(0:10)) +
    labs(x = "Phase",
         y = "Expectancy ratings (mean, SE)") +
  theme(legend.title = element_blank(),
        legend.position = c(0.9,0.91),
        panel.background = element_rect(fill = "white"),
        panel.grid.major = element_line(colour="grey", linetype = 3),
        panel.border = element_rect(colour = "black", fill = NA))
print(Figure_3a)
    
```

## Figure 3b
Produce single plot showing expectancy ratings by CS across the 3 phases of the experiment, by CS and by contingency awareness.  Use different shapes for each CS and different colours for the contingency awareness groups, and jitter the position of the error bars and their corresponding points and lines, so as to make them easier to distinguish.

```{r Fig3b}

# Add contingency awareness data to df_exp

df_exp %<>% mutate(awareness = case_when(
    id == 8 | id == 10 | id == 15 | id == 16 | id == 17 | id == 22 | id == 26 | id == 28 | id == 55 | id == 57 | id == 63 | id == 65 | id == 68 | id == 71 | id == 77 | id == 78 ~ "unaware",
    id == 12 | id == 14 | id == 19 | id == 24 | id == 25 | id == 27 | id == 29 | id == 30 | id == 31 | id == 33 | id == 35 | id == 40 | id == 41 | id == 42 | id == 43 | id == 45 | id == 46 | id == 50 | id == 56 | id == 58 | id == 59 | id == 61 | id == 64 | id == 69 | id == 70 | id == 73 | id == 76 ~ "aware"
))

check_missing <- df_exp %>% filter(is.na(awareness))
unique(check_missing$id)

# Prepare data frame

df_exp_plot3b <- df_exp %>% 
    group_by(awareness, phase, CS) %>%
    summarise(mean = mean(rating, na.rm = TRUE),
              se = sd(rating, na.rm = TRUE)/sqrt(n())) %>%
    filter(!is.na(awareness))
    
df_exp_plot3b %<>% mutate(plot_group = case_when(
    awareness == "aware" & CS == "CSplus" ~ "Aware group, CS+",
    awareness == "aware" & CS == "CSminus" ~ "Aware group, CS-",
    awareness == "unaware" & CS == "CSplus" ~ "Unaware group, CS+",
    awareness == "unaware" & CS == "CSminus" ~ "Unaware group, CS-"
))

df_exp_plot3b$phase <- factor(df_exp_plot$phase, levels = c("baseline", "acquisition", "test"))

Figure_3b <- ggplot(data = df_exp_plot3b) +
    aes(x = phase,
        y = mean,
        group = plot_group,
        colour = awareness,
        linetype = CS) +
    geom_point(position = position_dodge(0.1),
               aes(shape = CS),
               size = 4) +
    scale_color_manual(values=c("orange", "purple")) +
    geom_errorbar(aes(ymin = mean-se,
                      ymax = mean+se),
                  width = 0.2,
                  position = position_dodge(0.1)) +
    geom_line(position = position_dodge(0.1)) +
    scale_shape_discrete(solid = FALSE) +
    scale_linetype_manual(values = c("dashed", "solid")) +
    scale_y_continuous(limits = c(0,10),
                       breaks = c(0:10)) +
    labs(x = "Phase",
         y = "Expectancy ratings (mean, SE)") +
  theme(legend.title = element_blank(),
        legend.position = c(0.9,0.86),
        panel.background = element_rect(fill = "white"),
        panel.grid.major = element_line(colour="grey", linetype = 3),
        panel.border = element_rect(colour = "black", fill = NA))
print(Figure_3b)
```

# Plot intensity ratings

## Figure 3b
Exact mean intensity ratings (error bars = standard error) for the CS+/USP and CS-/USNP trials during acquisition and test phase.

```{r Fig4b}

# Prepare data frame

# Import data

df_ratings <- read_csv('/Users/tory/Google Drive/Git/CC_threshold/Juliane_all_data/data/R_data_complete_dataset_ratings') %>%
    # Remove irrelevant rows
    filter(trial_type == "CSplus_UShigh" | trial_type == "CSminus_USlow") %>%
    # Change phase labelling
    mutate(phase = case_when(
    block_number == "acq1" | block_number == "acq2" | block_number == "acq3" ~ "acquisition",
    block_number == "test1" | block_number == "test2" | block_number == "test3"  ~ "test"))

# Prepare data frame

df_plot4 <- df_ratings %>% 
    group_by(phase, trial_type) %>%
    summarise(mean = mean(rating, na.rm = TRUE),
              se = sd(rating, na.rm = TRUE)/sqrt(n()))

# Create plot

Figure_4 <- ggplot(data = df_plot4) +
    aes(x = phase,
        y = mean,
        group = trial_type,
        linetype = trial_type) +
    geom_point(aes(shape = trial_type),
               size = 4) +
    geom_errorbar(aes(ymin = mean-se,
                      ymax = mean+se),
                  width = 0.2) +
    geom_line() +
    scale_shape_discrete(solid = FALSE) +
    scale_linetype_manual(values = c("dashed", "solid")) +
    scale_y_continuous(limits = c(-50,50),
                       breaks = c(-50,-40,-30,-20,-10,0,10,20,30,40,50)) +
    geom_hline(yintercept = 0, linetype = 3) +
    labs(x = "Phase",
         y = "SPARS ratings (mean, SE)") +
  theme(legend.title = element_blank(),
        legend.position = c(0.87,0.91),
        panel.background = element_rect(fill = "white"),
        panel.grid.major = element_line(colour="grey", linetype = 3),
        panel.border = element_rect(colour = "black", fill = NA))
print(Figure_4)

```

## Session information

```{r session_info}
sessionInfo()
```