---
title: "Juliane's Masters study: plot trial ratings from all participants"
author: "Tory Madden"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output: 
  html_document:
    theme: yeti
    hightlight: pygments
    toc: true
    toc_depth: 5
    toc_float: true
    code_folding: show
---

```{r setup, include = FALSE}
# Load packages
library(readr)
library(tidyverse)
library(dplyr)
library(tidyr)
library(magrittr)
library(RColorBrewer)
library(ggplot2)

# knitr setup
knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE,
                      fig.height = 8,
                      fig.width = 8,
                      fig.path = './figures/rating_plots/')
```

```{r data, include = FALSE}

# Import data

df <- read.csv('/Users/tory/Google Drive/Git/CC_threshold/Juliane_all_data/data/R_data_complete_dataset_ratings')
```

# Visualise ratings

Here, we look at participant ratings in several different ways:
1. Ratings provided by each participant over the course of the procedure
2. Ratings provided by each participant of test (CS-UStest) trials only (therefore test phase only)
3. Ratings for each participant, looking at CS-UStest pairings as:
(a) delivered (ignoring any mis-localisation of CSs)
(b) perceived (considering how each CS was classified by the participant).
4. Ratings for all participants, looking at CS-UStest pairing as 
(a) delivered (ignoring any mis-localisation of CSs)
(b) perceived (considering how each CS was classified by the participant).

## 1. Ratings provided by each participant over the course of the procedure
Dashed vertical lines divide trials at the transitions between baseline, acquisition and test phases.
```{r whole_proc_plots, include = TRUE}

# Plot ratings

df_test <- df %>% filter(trial_type == "CSplus_UStest" | trial_type == "CSminus_UStest")

df_plot1 <- df %>% filter(id < 23)
df_plot2 <- df %>% filter(id > 23 & id < 42)
df_plot3 <- df %>% filter(id > 41 & id < 64)
df_plot4 <- df %>% filter(id > 64)

ggplot(data = df_plot1) +
  aes(x = trial_no,
      y = rating,
      colour = trial_type) +
  geom_point() +
  scale_colour_brewer(palette = "Set1") +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_vline(xintercept = 10.5, linetype = 3) +
  geom_vline(xintercept = 34.5, linetype = 3) +
  facet_wrap(~ id, ncol = 2) +
  labs(title = "Ratings over course of procedure",
       subtitle = "Faceted by participant ID")

ggplot(data = df_plot2) +
  aes(x = trial_no,
      y = rating,
      colour = trial_type) +
  geom_point() +
  scale_colour_brewer(palette = "Set1") +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_vline(xintercept = 10.5, linetype = 3) +
  geom_vline(xintercept = 34.5, linetype = 3) +
  facet_wrap(~ id, ncol = 2) +
  labs(title = "Ratings over course of procedure",
       subtitle = "Faceted by participant ID")

ggplot(data = df_plot3) +
  aes(x = trial_no,
      y = rating,
      colour = trial_type) +
  geom_point() +
  scale_colour_brewer(palette = "Set1") +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_vline(xintercept = 10.5, linetype = 3) +
  geom_vline(xintercept = 34.5, linetype = 3) +
  facet_wrap(~ id, ncol = 2) +
  labs(title = "Ratings over course of procedure",
       subtitle = "Faceted by participant ID")

ggplot(data = df_plot4) +
  aes(x = trial_no,
      y = rating,
      colour = trial_type) +
  geom_point() +
  scale_colour_brewer(palette = "Set1") +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_vline(xintercept = 10.5, linetype = 3) +
  geom_vline(xintercept = 34.5, linetype = 3) +
  facet_wrap(~ id, ncol = 2) +
  labs(title = "Ratings over course of procedure",
       subtitle = "Faceted by participant ID")
```

## 2. Ratings provided by each participant of test (CS-UStest) trials only (therefore test phase only)

```{r test_plots, include = TRUE}

df_test <- df %>% filter(trial_type == "CSplus_UStest" | trial_type == "CSminus_UStest")

dftest_plot1 <- df_test %>% filter(id < 23)
dftest_plot2 <- df_test %>% filter(id > 23 & id < 42)
dftest_plot3 <- df_test %>% filter(id > 41 & id < 64)
dftest_plot4 <- df_test %>% filter(id > 64)

ggplot(data = dftest_plot1) +
  aes(x = trial_no,
      y = rating,
      colour = trial_type) +
  geom_point() +
  scale_colour_brewer(palette = "Set1") +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_vline(xintercept = 70.5, linetype = 3) +
  geom_vline(xintercept = 34.5, linetype = 3) +
  geom_vline(xintercept = 105.5, linetype = 3) +
  facet_wrap(~ id, ncol =2) +
  labs(title = "Ratings over course of test phases, UStest trials only",
       subtitle = "Faceted by participant ID")

ggplot(data = dftest_plot2) +
  aes(x = trial_no,
      y = rating,
      colour = trial_type) +
  geom_point() +
  scale_colour_brewer(palette = "Set1") +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_vline(xintercept = 70.5, linetype = 3) +
  geom_vline(xintercept = 34.5, linetype = 3) +
  geom_vline(xintercept = 105.5, linetype = 3) +
  facet_wrap(~ id, ncol =2) +
  labs(title = "Ratings over course of test phases, UStest trials only",
       subtitle = "Faceted by participant ID")

ggplot(data = dftest_plot3) +
  aes(x = trial_no,
      y = rating,
      colour = trial_type) +
  geom_point() +
  scale_colour_brewer(palette = "Set1") +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_vline(xintercept = 70.5, linetype = 3) +
  geom_vline(xintercept = 34.5, linetype = 3) +
  geom_vline(xintercept = 105.5, linetype = 3) +
  facet_wrap(~ id, ncol =2) +
  labs(title = "Ratings over course of test phases, UStest trials only",
       subtitle = "Faceted by participant ID")

ggplot(data = dftest_plot4) +
  aes(x = trial_no,
      y = rating,
      colour = trial_type) +
  geom_point() +
  scale_colour_brewer(palette = "Set1") +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_vline(xintercept = 70.5, linetype = 3) +
  geom_vline(xintercept = 34.5, linetype = 3) +
  geom_vline(xintercept = 105.5, linetype = 3) +
  facet_wrap(~ id, ncol =2) +
  labs(title = "Ratings over course of test phases, UStest trials only",
       subtitle = "Faceted by participant ID")
```

## 3. Ratings for each participant, looking at CS-UStest pairings as:
(a) delivered (ignoring any mis-localisation of CSs)
(b) perceived (considering how each CS was classified by the participant).

```{r group_test_trials_actual}

ggplot(data = dftest_plot1) +
  aes(x = as.factor(trial_type),
      y = rating,
      colour = trial_type) +
  geom_jitter(height = 0) +
  geom_boxplot() +
  scale_colour_brewer(palette = "Set1") +
  facet_wrap(~ id) +
  labs(title = "Ratings of test stimuli according to actual trial type",
       subtitle = "Faceted by ID")

ggplot(data = dftest_plot2) +
  aes(x = as.factor(trial_type),
      y = rating,
      colour = trial_type) +
  geom_jitter(height = 0) +
  geom_boxplot() +
  scale_colour_brewer(palette = "Set1") +
  facet_wrap(~ id) +
  labs(title = "Ratings of test stimuli according to actual trial type",
       subtitle = "Faceted by ID")

ggplot(data = dftest_plot3) +
  aes(x = as.factor(trial_type),
      y = rating,
      colour = trial_type) +
  geom_jitter(height = 0) +
  geom_boxplot() +
  scale_colour_brewer(palette = "Set1") +
  facet_wrap(~ id) +
  labs(title = "Ratings of test stimuli according to actual trial type",
       subtitle = "Faceted by ID")

ggplot(data = dftest_plot4) +
  aes(x = as.factor(trial_type),
      y = rating,
      colour = trial_type) +
  geom_jitter(height = 0) +
  geom_boxplot() +
  scale_colour_brewer(palette = "Set1") +
  facet_wrap(~ id) +
  labs(title = "Ratings of test stimuli according to actual trial type",
       subtitle = "Faceted by ID")

ggplot(data = dftest_plot1) +
  aes(x = as.factor(perc_trial_type),
      y = rating,
      colour = perc_trial_type) +
  geom_jitter(height = 0) +
  geom_boxplot() +
  scale_colour_brewer(palette = "Set1") +
  facet_wrap(~ id) +
  labs(title = "Ratings of test stimuli according to perceived trial type",
       subtitle = "Faceted by ID")

ggplot(data = dftest_plot2) +
  aes(x = as.factor(perc_trial_type),
      y = rating,
      colour = perc_trial_type) +
  geom_jitter(height = 0) +
  geom_boxplot() +
  scale_colour_brewer(palette = "Set1") +
  facet_wrap(~ id) +
  labs(title = "Ratings of test stimuli according to perceived trial type",
       subtitle = "Faceted by ID")

ggplot(data = dftest_plot3) +
  aes(x = as.factor(perc_trial_type),
      y = rating,
      colour = perc_trial_type) +
  geom_jitter(height = 0) +
  geom_boxplot() +
  scale_colour_brewer(palette = "Set1") +
  facet_wrap(~ id) +
  labs(title = "Ratings of test stimuli according to perceived trial type",
       subtitle = "Faceted by ID")

ggplot(data = dftest_plot4) +
  aes(x = as.factor(perc_trial_type),
      y = rating,
      colour = perc_trial_type) +
  geom_jitter(height = 0) +
  geom_boxplot() +
  scale_colour_brewer(palette = "Set1") +
  facet_wrap(~ id) +
  labs(title = "Ratings of test stimuli according to perceived trial type",
       subtitle = "Faceted by ID")
```

4. Ratings for all participants, looking at CS-UStest pairing as 
(a) delivered (ignoring any mis-localisation of CSs)
(b) perceived (considering how each CS was classified by the participant).

```{r }
df_test %>% filter(!is.na(trial_type)) %>%
ggplot(data = .) +
  aes(x = as.factor(trial_type),
      y = rating,
      colour = trial_type) +
  geom_jitter(height = 0) +
  geom_boxplot() +
  scale_colour_brewer(palette = "Set1") +
  labs(title = "Ratings of test stimuli",
       subtitle = "According to actual trial type")

df_test %>% filter(!is.na(perc_trial_type)) %>%
  ggplot(data = .) +
  aes(x = as.factor(perc_trial_type),
      y = rating,
      colour = perc_trial_type) +
  geom_jitter(height = 0) +
  geom_boxplot() +
  scale_colour_brewer(palette = "Set1") +
  labs(title = "Ratings of test stimuli",
       subtitle = "According to perceived trial type")
```

# Prediction intervals

Here, we visualise the data using prediction intervals.  This is a statistically calculated range in which the rating of each CSminus-UStest is predicted to fall.  The question of interest is whether or not any CS+-UStest trials fall **outside** and **to the right of** that prediction interval.  If they do, it would indicate a statistical likelihood that that particular trial was experienced as more intense than might be expected on the basis of chance. 

In order to assist with visualisation, trials that do fall **outside** the prediction interval are coloured BLUE.

## Test trials
First, we consider test trials as delivered (i.e. ignoring any mis-localisation by the participant).  Second, we consider test trials as perceived.  

```{r prediction_delivered}

# Prediction intervals for test trials as DELIVERED
df_prediction <- df_test %>%
  # Filter for CSminus_UStest
  filter(trial_type == 'CSminus_UStest' & id != 23) %>%
  # Calculate 95% prediction interval 
  group_by(id) %>%
  dplyr::summarise(lower_pi = mean(rating, na.rm = TRUE) - 1.96 * (sd(rating, na.rm = TRUE)),
            upper_pi = mean(rating, na.rm = TRUE) + 1.96 * (sd(rating, na.rm = TRUE))) %>%
  # Truncate PIs at +50 and -50
  mutate(lower_pi = ifelse(lower_pi < -50,
                           yes = -50,
                           no = lower_pi),
         upper_pi = ifelse(upper_pi > 50,
                           yes = 50,
                           no = upper_pi)) %>%
  # Join back with df_test
  left_join(df_test) %>%
  # Is CSplus_UStest within the prediction interval
  mutate(CC_effect = ifelse(rating > upper_pi,
                               yes = 'yes',
                               no = 'no')) %>%
  # Filter for CSplus_UStest
  filter(trial_type == 'CSplus_UStest') 

ggplot(data = df_prediction) +
  aes(x = as.factor(id),
      y = rating) +
  geom_jitter(aes(colour = CC_effect),
             height = 0) +
  geom_errorbar(aes(ymin = lower_pi,
                    ymax = upper_pi)) +
  ylim(-50, 50) +
  geom_hline(yintercept = 0, linetype = 2) +
  scale_colour_brewer(palette = "Set1") +
  coord_flip() +
  labs(title = 'Prediction interval for CSminus_UStest trial ratings',
       subtitle = 'Points are individual CSplus_UStest trial ratings',
       y = 'FEST rating',
       x = 'Participant ID')

# Prediction intervals for test trials as PERCEIVED

df_prediction_perc <- df_test %>%
  filter(!is.na(perc_trial_type)) %>%
  # Filter for CSminus_UStest
  filter(perc_trial_type == 'CSminus_UStest' & id != 23 & id != 19) %>%
  # Calculate 95% prediction interval 
  group_by(id) %>%
  summarise(lower_pi = mean(rating, na.rm = TRUE) - 1.96 * (sd(rating, na.rm = TRUE)),
            upper_pi = mean(rating, na.rm = TRUE) + 1.96 * (sd(rating, na.rm = TRUE))) %>%
  # Truncate PIs at +50 and -50
  mutate(lower_pi = ifelse(lower_pi < -50,
                           yes = -50,
                           no = lower_pi),
         upper_pi = ifelse(upper_pi > 50,
                           yes = 50,
                           no = upper_pi)) %>%
  # Join back with df_test
  left_join(df_test) %>%
  # Is CSplus_UStest within the prediction interval
  mutate(CC_effect = ifelse(rating > upper_pi,
                            yes = 'yes',
                            no = 'no')) %>%
  # Filter for CSplus_UStest
  filter(perc_trial_type == 'CSplus_UStest') 

# Save file df_prediction_perc

write_rds(x = df_prediction_perc,
          path = '/Users/tory/Google Drive/Git/CC_threshold/Juliane_all_data/data/Juliane_raw_as_perc.rds')

ggplot(data = df_prediction_perc) +
  aes(x = as.factor(id),
      y = rating) +
  geom_jitter(aes(colour = CC_effect),
              height = 0) +
  geom_errorbar(aes(ymin = lower_pi,
                    ymax = upper_pi)) +
  ylim(-50, 50) +
  geom_hline(yintercept = 0, linetype = 2) +
  scale_colour_brewer(palette = "Set1") +
  coord_flip() +
  labs(title = 'Prediction interval for ratings of trials perceived as CSminus_UStest ',
       subtitle = 'Points are individual CSplus_UStest (percieved) trial ratings',
       y = 'FEST rating',
       x = 'Participant ID')
```

## Reinforced trials
Now we consider prediction intervals for all reinforced (CSplus-UShigh and CSminus-USlow) trials.  We want to check that there are plenty of CSplus-UShigh trials that fall outside the prediction interval for the CSminus-USlow trials, in order to confirm that these two trial types were actually experienced differently. We also want the prediction intervals for CSminus-USlow trials NOT to cross zero (i.e. they were reliably non-painful).

```{r prediction_perceived}

# Prediction intervals for all reinforced (CS+/USH or CS-/USL) trials (as delivered)

df_prediction_reinf <- df %>%
  # Filter for CSminus_USlow
  filter(trial_type == 'CSminus_USlow' & id != 23) %>%
  # Calculate 95% prediction interval 
  group_by(id) %>%
  dplyr::summarise(lower_pi = mean(rating, na.rm = TRUE) - 1.96 * (sd(rating, na.rm = TRUE)),
            upper_pi = mean(rating, na.rm = TRUE) + 1.96 * (sd(rating, na.rm = TRUE))) %>%
  # Truncate PIs at +50 and -50
  mutate(lower_pi = ifelse(lower_pi < -50,
                           yes = -50,
                           no = lower_pi),
         upper_pi = ifelse(upper_pi > 50,
                           yes = 50,
                           no = upper_pi)) %>%
  # Join back with df_test
  left_join(df) %>%
  # Is CSplus_UShigh within the prediction interval
  mutate(diff_intensities = ifelse(rating > upper_pi,
                            yes = 'yes',
                            no = 'no')) %>%
  # Filter for CSplus_UShigh
  filter(trial_type == 'CSplus_UShigh') 

ggplot(data = df_prediction_reinf) +
  aes(x = as.factor(id),
      y = rating) +
  geom_jitter(aes(colour = diff_intensities),
              height = 0) +
  geom_errorbar(aes(ymin = lower_pi,
                    ymax = upper_pi)) +
  ylim(-50, 50) +
  geom_hline(yintercept = 0, linetype = 2) +
  scale_colour_brewer(palette = "Set1") +
  coord_flip() +
  labs(title = 'Prediction interval for CSminus_USlow trial ratings',
       subtitle = 'Points are individual CSplus_UShigh trial ratings',
       y = 'FEST rating',
       x = 'Participant ID')
```

### Interpretation: 
Calibration was poor.  However, participants 11, 15, 26, 33, 50, 58, 61 and 78 have a prediction interval for their CS-/USlow trials that does not cross zero - i.e. good calibration.  Check for CC effect in above graphs.

## Session information

```{r session_info}
sessionInfo()
```