---
title: "Juliane's Masters study: run a MH test on data from all participants"
author: "Tory Madden"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output: 
  html_document:
    theme: yeti
    hightlight: pygments
    toc: true
    toc_depth: 5
    toc_float: true
    code_folding: show
---

```{r setup, include = FALSE}
# Load packages
library(vcdExtra)
library(tidyverse)

# knitr setup
knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE,
                      fig.height = 8,
                      fig.width = 8,
                      fig.path = './figures/expectancy_plots/')
```

```{r data, include = FALSE}

# Import and prepare data


df <- read_csv('/Users/tory/Google Drive/Git/CC_threshold/Juliane_all_data/data/R_data_complete_dataset_ratings')

df %<>% mutate(painful = ifelse(rating > 0, 
                                yes = "yes",
                                no = "no")) %>%
  filter(trial_type == "CSplus_UStest" | trial_type == "CSminus_UStest")
```

# Cross-tabulation table

```{r crosstabs_table}
df_tab <- xtabs(~ trial_type + painful + id, data = df)

# Look at table
ftable(df_tab)
```

# Mantel Haenszel test

We can use a Mantel-Haenszel test because we don't need to allow for ordered variables (e.g. very sick, slightly sick, not sick).  This test stratifies by ID before it pools all the IDs.  It also applied a contoinuity correction in its default form, which correct the p-value over-estimation that tends to happen in a chi square with 2 x 2 arrangement - therefore this version is more conservative than the CMH option considered below.  

```{r MH_test}

mantelhaen.test(df_tab)
```

### Interpretation
This result indicates a meaningful difference.

# Mosaic plot
Check the direction of the effect by creating a mosaic plot:

```{r mosaic_plot}

# First remove the stratification.

df_tab2 <- xtabs(~ trial_type + painful, data = df)

mosaicplot(df_tab2)
```

### Note:
If we wanted to add other stratification (e.g. test for effect of test phse block, we could run a CMH test (which allows for ordinal variables as well as stratification).
We would do that by using the syntax
CMHtest(df_tab, overall = TRUE)
Currently, this produces an error - possibly becasue we haven't provided an ordinal variable.

## Session information

```{r session_info}
sessionInfo()
```